---
title: "Examples"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Worked examples for the tasks CBOR is usually used for. Each one runs as
written, with only zucbor and base R.

```{r}
library(zucbor)

# CBOR examples are usually written in hex; this turns hex into bytes.
hex <- function(h) {
  h <- gsub("[^0-9a-fA-F]", "", h)
  as.raw(strtoi(substring(h, seq(1, nchar(h), 2), seq(2, nchar(h), 2)), 16L))
}
```

## Look at bytes before decoding them

Diagnostic notation is the quickest way to see what a message holds. It
checks the input first, so it is safe on anything:

```{r}
msg <- hex("a3 63 74 6d 70 f9 4d 70 62 69 64 1a 00 01 e2 40 62 6f 6b f5")
cbor_diagnose(msg)
cbor_validate(msg)
str(cbor_decode(msg))
```

A malformed message says what is wrong and where:

```{r}
bad <- hex("a3 63 74 6d 70 f9 4d 70 62 69 64")    # cut off mid-way
cbor_validate(bad)
tryCatch(cbor_validate(bad, error = TRUE),
         zucbor_error = function(e) c(class = class(e)[1], offset = e$offset))
```

## Maps with integer keys

Protocols built on CBOR save space by using small integers as map keys.
Such a map decodes to a `cbor_map`, which keeps the keys as they are. A
small helper looks values up by key:

```{r}
cose_key <- cbor_decode(hex(paste0(
  "a5 01 02 03 26 20 01",
  "21 58 20 65eda5a12577c2bae829437fe338701a10aaa375e1bb5b5de108de439c08551d",
  "22 58 20 1e52ed75701163f7f9e40ddf9f341b3dc9ba860af7e0ca7ca7e9eecd0084d19c")))
cose_key

get <- function(map, key) {
  m <- unclass(map)
  hit <- vapply(m$keys, identical, logical(1), key)
  if (any(hit)) m$values[[which(hit)[1]]] else NULL
}
get(cose_key, 3L)      # alg: -7, ES256
get(cose_key, -1L)     # crv: 1, P-256
```

`map_keys = "string"` names every entry instead, turning non-text keys into
their diagnostic notation. It is handy for exploring, but lossy: `1` and
`"1"` would collide, and zucbor refuses that rather than drop one.

```{r}
names(cbor_decode(hex("a3 01 02 03 26 20 01"), map_keys = "string"))
```

To write such a map, build it with `cbor_map()`. Whole numbers are written
as integers even when R holds them as doubles, as the protocol needs:

```{r}
hdr <- cbor_map(list(1L, 4L), list(-7, charToRaw("key-1")))
cbor_diagnose(cbor_encode(hdr))
```

## Integers beyond 2^53

A double holds integers exactly only up to 2^53. Wider ones, such as
64-bit identifiers or nanosecond timestamps, decode to `cbor_bigint`, which
keeps every digit:

```{r}
id <- cbor_decode(hex("1b 16 f8 44 1f 3b 6f 00 01"))
id
as.character(id)
cbor_encode(id)                           # the same nine bytes back
cbor_decode(hex("1b 16 f8 44 1f 3b 6f 00 01"), big_integers = "double")
```

## Dates and times

Tags 0 and 1 are instants and decode to `POSIXct` in UTC; tags 100 and
1004 are calendar days and decode to `Date`. `POSIXct` is written back as
tag 1, `Date` as tag 1004:

```{r}
cbor_decode(hex("c0 74 32 30 31 33 2d 30 33 2d 32 31 54 32 30 3a 30 34 3a 30 30 5a"))
cbor_decode(hex("c1 1a 51 4b 67 b0"))
when <- as.POSIXct("2026-09-30 12:00:00", tz = "UTC")
cbor_diagnose(cbor_encode(list(seen = when, day = as.Date(when))))
```

Other tags come back as `cbor_tag`, with the number and the content:

```{r}
uri <- cbor_decode(hex("d8 20 76 68 74 74 70 3a 2f 2f 77 77 77 2e 65 78 61 6d 70 6c 65 2e 63 6f 6d"))
uri$tag
uri$value
```

## Tags of your own

zucbor converts a handful of tags; the rest come back as `cbor_tag`. Two
hooks give any other tag a meaning in R, without zucbor knowing about it:
`tag_handlers` when decoding, and an `as_cbor()` method when encoding.
Together they make a class of your own round-trip. Handlers run only after
the whole input has passed its checks, so they never see malformed bytes.

### UUIDs (tag 37)

RFC 9562 UUIDs travel as tag 37 around their 16 bytes. Here they become a
small class holding the usual text form:

```{r}
uuid <- function(text) structure(tolower(text), class = "uuid")
format.uuid <- function(x, ...) unclass(x)
print.uuid <- function(x, ...) print(paste0("<uuid ", format(x), ">"), quote = FALSE)

as_cbor.uuid <- function(x, ...) {
  h <- gsub("-", "", unclass(x))
  cbor_tag(37, as.raw(strtoi(substring(h, seq(1, 31, 2), seq(2, 32, 2)), 16L)))
}
uuid_from_bytes <- function(bytes) {
  # The content is checked CBOR, but not checked to be a UUID: that is ours.
  if (!is.raw(bytes) || length(bytes) != 16L) stop("a UUID is 16 bytes")
  h <- paste(sprintf("%02x", as.integer(bytes)), collapse = "")
  uuid(paste(substring(h, c(1, 9, 13, 17, 21), c(8, 12, 16, 20, 32)), collapse = "-"))
}

id <- uuid("F81D4FAE-7DEC-11D0-A765-00A0C91E6BF6")
bytes <- cbor_encode(list(id = id, n = 1L))
cbor_diagnose(bytes)

cbor_decode(bytes, tag_handlers = list("37" = uuid_from_bytes))
```

A method defined in a script is found from the global environment. A
package registers it instead, with `S3method(zucbor::as_cbor, uuid)` in its
`NAMESPACE` (roxygen: `@exportS3Method zucbor::as_cbor`).

### IP addresses (tags 52 and 54)

RFC 9164 writes an IPv4 address as tag 52 around 4 bytes and an IPv6
address as tag 54 around 16, and a prefix as an array of its length and
the address bytes with trailing zero bytes left out. Decoding them to text
takes two handlers:

```{r}
ip_text <- function(v, width) {
  prefix <- NULL
  if (is.list(v)) {                              # [length, bytes]: a prefix
    prefix <- v[[1]]
    v <- v[[2]]
  }
  b <- as.integer(c(v, raw(width - length(v))))  # put back the zero bytes
  addr <- if (width == 4) {
    paste(b, collapse = ".")
  } else {
    paste(sprintf("%x", b[c(TRUE, FALSE)] * 256L + b[c(FALSE, TRUE)]), collapse = ":")
  }
  if (is.null(prefix)) addr else paste0(addr, "/", prefix)
}
ip <- list("52" = function(v) ip_text(v, 4), "54" = function(v) ip_text(v, 16))

cbor_decode(hex("d8 34 44 c0 00 02 01"), tag_handlers = ip)            # 192.0.2.1
cbor_decode(hex("d8 34 82 18 18 43 c0 00 02"), tag_handlers = ip)      # 192.0.2.0/24
cbor_decode(hex("d8 36 50 20 01 0d b8 12 34 de ed be ef ca fe fa ce fe ed"), tag_handlers = ip)
```

### Decimal fractions (tag 4)

Tag 4 is `[exponent, mantissa]`, the value mantissa × 10^exponent, so
273.15 is `[-2, 27315]` exactly. The [decimal](https://pedrobtz.github.io/decimal/)
package holds such numbers exactly in R, and stores each as its text,
which makes both directions a few lines of string handling. The mantissa
may be wider than 2^53; `cbor_bigint()` writes it exactly either way.

```{r, eval = requireNamespace("decimal", quietly = TRUE)}
as_cbor.decimal <- function(x, ...) {
  one <- function(s) {
    m <- regmatches(s, regexec("^(-?)([0-9]*)[.]?([0-9]*)(?:[eE]([-+]?[0-9]+))?$", s, perl = TRUE))[[1]]
    if (length(m) == 0L) stop("no CBOR form for decimal ", s)
    digits <- sub("^0+(?=.)", "", paste0(m[3], m[4]), perl = TRUE)
    exponent <- (if (nzchar(m[5])) as.integer(m[5]) else 0L) - nchar(m[4])
    cbor_tag(4, list(exponent, cbor_bigint(paste0(if (digits != "0") m[2], digits))))
  }
  s <- as.character(x)
  if (length(s) == 1L) one(s) else lapply(s, one)
}
decimal_from_tag <- function(v) {
  digits <- if (inherits(v, "cbor_bigint")) as.character(v) else sprintf("%.0f", as.numeric(v))
  decimal::decimal(paste0(digits[2], "E", digits[1]))
}

price <- decimal::decimal("273.15")
bytes <- cbor_encode(price)
bytes
cbor_diagnose(bytes)
cbor_decode(bytes, tag_handlers = list("4" = decimal_from_tag))
```

### Errors in handlers

Validity of the CBOR is zucbor's to check; what the content means is the
handler's. A handler that fails, as `uuid_from_bytes()` does on two bytes,
raises `zucbor_handler_error`, which carries the tag number and the
handler's own error:

```{r}
e <- tryCatch(
  cbor_decode(hex("d8 25 42 01 02"), tag_handlers = list("37" = uuid_from_bytes)),
  zucbor_handler_error = function(e) e
)
e$tag
conditionMessage(e$parent)
```

## CBOR inside CBOR

COSE and CWT carry CBOR inside byte strings, and tag 24 marks embedded
CBOR. zucbor never decodes those on its own; decoding them is one explicit
call, with its own checks and limits:

```{r}
outer <- cbor_decode(hex("d8 18 45 82 01 61 61 00"))
outer
```

That tagged byte string is five bytes of CBOR followed by one more byte,
so `cbor_decode()` refuses it, `cbor_decode_seq()` reads both items, and
`cbor_decode_prefix()` reads the first and says where it ended:

```{r}
inner <- hex("82 01 61 61 00")
tryCatch(cbor_decode(inner), zucbor_error = function(e) class(e)[1])
cbor_decode_seq(inner)
cbor_decode_prefix(inner)$consumed
```

When the embedded item is meant to be decoded, a handler for tag 24 does it
in place. It is still an explicit call, so it can set limits of its own:

```{r}
cbor_decode(hex("d8 18 44 82 01 61 61"),
            tag_handlers = list("24" = function(b) cbor_decode(b, max_depth = 4)))
```

## Telemetry as a data frame

Sensors often send records one after another as an RFC 8742 sequence.
Records with the same text keys become named lists, and a data frame is
one step away:

```{r}
readings <- lapply(1:5, function(i) list(sensor = paste0("t", i %% 2), c = 20 + i / 4, ok = i != 3))
stream <- cbor_encode_seq(readings)
length(stream)

records <- cbor_decode_seq(stream)
do.call(rbind, lapply(records, as.data.frame))
```

The columns come back as `c`, `ok`, `sensor`, not in the order the lists
were built: deterministic encoding writes map keys sorted by their encoded
bytes, shorter keys first, and decoding keeps the order it reads. Select
columns by name when order matters.

## Numeric data as typed arrays

RFC 8746 writes a numeric vector as one byte string instead of one item
per element. zucbor always decodes such typed arrays, and writes them when
asked, along with matrices as tag 1040:

```{r}
m <- matrix(c(1.5, NA, -0, Inf), 2)
bytes <- cbor_encode(m, typed_arrays = TRUE)
cbor_diagnose(bytes)
identical(cbor_decode(bytes), m)
```

Every bit survives, `NA` and `-0` included. For large data it is smaller and
much faster in both directions:

```{r}
x <- runif(1e5)
c(array = length(cbor_encode(x)), typed = length(cbor_encode(x, typed_arrays = TRUE)))
```

It is off by default because many CBOR decoders do not implement RFC 8746;
use it when you know the reader does.

## Files and connections

`cbor_read()` reads a file, a URL or a connection, and never reads more
than `max_size + 1` bytes, so a huge or endless source fails quickly:

```{r}
path <- tempfile(fileext = ".cbor")
writeBin(cbor_encode(list(version = 2L, tags = c("a", "b"))), path)
cbor_read(path)     # keys in encoded order: "tags" is shorter than "version"

tryCatch(cbor_read(path, max_size = 8), zucbor_limit_error = function(e) e$limit)
unlink(path)
```

An open connection is read from its current position and left open, so a
header can be read by hand first:

```{r}
con <- rawConnection(c(charToRaw("CBOR"), cbor_encode(1:3)))
readBin(con, "raw", n = 4)        # a 4-byte header, say
cbor_read(con)
close(con)
```

## Limits for untrusted input

The defaults suit messages from a network peer. Tighten them for small,
known messages, or loosen them for large trusted files:

```{r}
nested <- c(rep(as.raw(0x81), 20), as.raw(0x00))    # 20 nested arrays
tryCatch(cbor_decode(nested, max_depth = 8), zucbor_limit_error = function(e) conditionMessage(e))

big <- cbor_encode(as.list(1:5000))
tryCatch(cbor_decode(big, max_items = 1000), zucbor_limit_error = function(e) class(e)[1])
length(cbor_decode(big, max_items = Inf))
```

Duplicate map keys are refused by default, compared by value, so a key
written twice in different lengths is still caught:

```{r}
dup <- hex("a2 01 00 18 01 00")      # {1: 0, 1: 0}, the second key two bytes long
tryCatch(cbor_decode(dup), zucbor_duplicate_key = function(e) conditionMessage(e))
```

## Bytes that must not change

A signature covers bytes. Deterministic encoding means that decoding a
deterministic message and encoding it again gives exactly those bytes, so
a signed structure can be taken apart, checked and rebuilt safely:

```{r}
signed <- hex("84 43 a1 01 26 a0 45 68 65 6c 6c 6f 40")
cbor_validate(signed, deterministic = TRUE)
identical(cbor_encode(cbor_decode(signed)), signed)
```

Order does not leak into the bytes: two lists with the same entries encode
identically, whatever order they were built in.

```{r}
identical(cbor_encode(list(b = 2, a = 1)), cbor_encode(list(a = 1, b = 2)))
```

A one-element array comes back marked with `I()`, which is what keeps its
brackets when it is encoded again:

```{r}
one <- cbor_decode(hex("a1 63 78 35 63 81 42 01 02"))    # {"x5c": [h'0102']}
str(one)
cbor_encode(one)
```
