---
title: "Getting started with zucbor"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r}
library(zucbor)
```

## A round trip

```{r}
x <- list(name = "sensor-7", reading = c(21.5, 21.75), ok = TRUE,
          seen = as.POSIXct("2026-09-30 12:00:00", tz = "UTC"))
b <- cbor_encode(x)
length(b)
cbor_diagnose(b)
str(cbor_decode(b))
```

## How CBOR becomes R

| CBOR | R |
|---|---|
| integer | `integer`, `double` up to 2^53, `cbor_bigint` beyond |
| float | `double` |
| `true`, `false` | `logical` |
| `null`, `undefined` | `NULL`, or `NA` in a vector |
| byte string | `raw` |
| text string | `character` |
| array | an atomic vector when the elements agree, else a `list` |
| map with text keys | a named `list` |
| any other map | `cbor_map` |
| date and time tags | `POSIXct`, `Date` |
| other tags | `cbor_tag` |

```{r}
cbor_decode(cbor_encode(list(1L, "a")))           # mixed: a list
cbor_decode(as.raw(c(0x82, 0x01, 0x02)))          # [1, 2]: an integer vector
cbor_decode(as.raw(c(0x81, 0x01)))                # [1]: I(1L), so it re-encodes as [1]
cbor_decode(as.raw(c(0x1b, rep(0xff, 8))))        # 2^64 - 1, exactly
```

## Values R has no type for

```{r}
cbor_encode(cbor_map(list(1L, -1L), list("a", "b")))    # integer keys
cbor_encode(cbor_tag(32, "https://example.com"))         # a URI tag
cbor_encode(cbor_bigint("340282366920938463463374607431768211456"))   # 2^128
```

## Sequences

An RFC 8742 CBOR sequence is items one after another, as sensors stream them:

```{r}
s <- cbor_encode_seq(list(1, "two", list(three = 3)))
cbor_decode_seq(s)
```
